\clearpage \section{Methodology}

Since the essential part of this work is the creation of various machine learning models and the procurement of the data to create them, this work reaches into the field of \emph{Data Mining}. For this reason, the implementation method is based on an established procedure in this field: the \emph{CRISP-DM} \parencite{chapman_crisp-dm_2000} procedure. This process builds the foundation of \citeauthor[]{geron_hands-machine_2019}'s data mining process presented in his book '\citetitle{geron_hands-machine_2019}', which offers a systematic approach to the type of problem at hand. This methodology is therefore used to systematically follow the data procurement process and the creation and evaluation of the different models.

Aligned with standard research designs, two types of models emerge in the data mining process. The first type of model is the \emph{baseline} model, which will mainly use historical data from the Fantasy League provider for its predictions. The second type is the \emph{treatment} model, which will additionally have access to the betting odds. With this design, a \emph{t}-test can be used in the end to test whether betting odds have a statistically significant effect on the predictions. In parallel, Cohen's D is calculated and used to determine the effect size.

The resulting models are compared on a test dataset experimentally. The experiment will pitch the baseline models against the treatment models and investigate whether the treatment models actually create better teams and thus perform better. For this purpose, standard metrics for the evaluation of machine learning models are used. Moreover, additional benchmarks are calculated, such as the optimal possible team or the rank in the player ranking, to investigate how such models perform against real players.

\clearpage In this paper, reference is made to previous work in the literature. The current state of research is determined with the help of a literature review, which is conducted according to the established standards of \citeauthor[]{webster_guest_2002} and \citeauthor[]{vom_brocke_standing_2015}, to ensure comprehensibly and reproducibility.

Following the methodologies above, the paper is divided into three main chapters. First, the basics of fantasy leagues are explained and then specified in more detail for the chosen provider \emph{SPITCH}. Subsequently, the rules of \emph{SPITCH} are explained shortly, limited to the essentials. In the next chapter, the literature review takes place, which presents the current state of research. In the final chapter, the data mining process takes place, where data is collected, processed and analysed, and models are created, optimised and compared with each other. This process results in two final models: a baseline and a treatment model. Finally, these two models are tested against each other in an experiment, and the effect size of betting odds is measured and evaluated.
